\documentclass{beamer}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{tikz}
\usetheme{Madrid}

\title{Rethinking Agent-Tool Integration}
\subtitle{Beyond Middlemen: A Case for Direct, Descriptive Protocols}
\author{}
\date{June 2025}

\begin{document}

% Title slide
\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{What is the Model Context Protocol (MCP)?}
  \centering\large\textit{The ``USB-C for AI Agents''}
  \vfill
  \begin{itemize}
    \item A protocol designed to standardize how AI agents communicate with external tools.
    \vfill
    \item It acts as a \textbf{translation layer} or \textbf{middleman}. An MCP server ``wraps'' an existing tool (e.g., a REST API) and exposes it in a model-friendly format.
    \vfill
    \item \textbf{The Goal:} Allow any agent to connect to any tool without writing custom ``glue code'' for each integration.
  \end{itemize}
\end{frame}

\begin{frame}{The MCP Interaction Loop}
  How an agent calls a tool via MCP:
  \begin{enumerate}
    \item<1-> The \textbf{Agent (Host)} needs to use a tool.
    \vfill
    \item<2-> It sends a request to a dedicated \textbf{MCP Server} that wraps the tool.
    \vfill
    \item<3-> The MCP Server translates the request and calls the \textbf{native tool API}.
    \vfill
    \item<4-> The Server translates the native response back into the MCP format.
    \vfill
    \item<5-> The Agent receives the standardized response.
  \end{enumerate}
\end{frame}

% Introduction: Two Philosophies
\begin{frame}{The Core Question: Middleman or Manual?}
  How should an agent interact with an external tool?
  \vfill
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      \centering
      \textbf{The Middleman (MCP)}
      \begin{itemize}
        \item Forces all traffic through a new protocol layer.
        \item Requires writing ``wrappers'' for existing tools.
        \item Must reinvent solutions for auth, security, etc.
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centering
      \textbf{The Manual (UTCP)}
      \begin{itemize}
        \item Describes how to call a tool's \textit{native} interface.
        \item Gets out of the way.
        \item Leverages existing, battle-tested infrastructure.
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

% Section: The Problem with Middlemen
\section{The Problem with the Middleman Approach}

\begin{frame}{MCP's Architecture: The Mandatory Middleman}
  \begin{itemize}
    \item MCP requires that every tool be placed behind a running MCP server.
    \item All agent requests are proxied through this server.
    \item This architecture creates a fundamental problem: it forces developers to \textbf{reinvent the wheel}.
  \end{itemize}
  \vfill
  \begin{beamercolorbox}[sep=1em,wd=\textwidth]{alertblock}
    Instead of using a tool's native endpoint, you must build and maintain a new service just to translate calls.
  \end{beamercolorbox}
\end{frame}

\begin{frame}{Symptom 1: The "Wrapper Tax"}
  The most immediate cost of the middleman is the ``wrapper tax'':
  \begin{itemize}
    \item \textbf{Protocol Mismatch:} A REST API needs a wrapper to speak MCP's HTTP streaming. A CLI tool needs a wrapper to be a live server. This is redundant engineering effort.
    \item \textbf{Infrastructure Overhead:} Every tool now requires a running process, consuming resources and adding a point of failure.
    \item \textbf{No Native Integration:} You can't just point to an existing gRPC service, WebSocket, or serverless function. It must be adapted.
  \end{itemize}
\end{frame}

\begin{frame}{Symptom 2: Re-implementing Core Services}
  Because the middleman isolates the agent from the native tool, it must re-solve problems the native tool already solved:
  \begin{itemize}
    \item \textbf{Authentication:} How does the MCP server authenticate to the real API? It needs its own logic, separate from the agent's credentials.
    \item \textbf{Security & Permissions:} The native tool has roles and permissions. The MCP wrapper must now re-implement or proxy this logic, adding complexity and risk.
    \item \textbf{Billing & Rate Limiting:} A paid API's billing is tied to its native endpoint. Routing through MCP obscures usage and makes cost tracking difficult.
  \end{itemize}
\end{frame}

\begin{frame}{Symptom 3: Inefficiency and Opacity}
  Forcing everything through a proxy layer is inherently inefficient:
  \begin{itemize}
    \item \textbf{Context Overload:} The agent's context is filled with MCP-specific metadata, not just the essential tool information.
    \item \textbf{Latency:} Every call involves an extra network hop (Agent -> MCP Server -> Tool).
    \item \textbf{Unstructured Data:} The middleman often strips rich, structured data from the native tool, passing along plain text that the agent must re-parse.
  \end{itemize}
\end{frame}

\begin{frame}{Symptom 4: Brittle Standard, Slow Evolution}
  The middleman design makes the protocol itself difficult to evolve:
  \begin{itemize}
    \item \textbf{Tightly Coupled Transport:} MCP's logic is tied to its transport (e.g., HTTP streaming). When the standard changes (e.g., SSE to HTTP streaming), it's a \textit{breaking change for everyone}.
    \vfill
    \item \textbf{Ecosystem-Wide Rebuild:} 
    \begin{itemize}
        \item Client builders must support new and legacy versions.
        \item Server providers must rebuild and redeploy their wrappers.
    \end{itemize}
    \vfill
    \item \textbf{Result:} This high cost of change creates inertia, slowing down the adoption of improvements and bug fixes for the protocol itself.
  \end{itemize}
\end{frame}

% Section: A Better Way
\section{A Better Way: The "Manual" Protocol}

\begin{frame}{Introducing UTCP: A Manual for Tools}
  UTCP is designed with a simple philosophy: a tool-calling protocol should be a descriptive manual, not a prescriptive middleman.
  \begin{tcolorbox}[colback=green!5!white,colframe=green!75!black,title=The UTCP Philosophy]
    A UTCP definition simply tells an agent: ``Here is a tool. Here is its \textbf{native endpoint} (HTTP, gRPC, CLI, etc.), and here is \textbf{how to call it directly}.''
  \end{tcolorbox}
  \vfill
  \begin{itemize}
    \item The protocol gets out of the way after discovery.
    \item The agent communicates directly with the tool.
  \end{itemize}
\end{frame}

\begin{frame}{How the "Manual" Solves the Core Problem}
  By telling the agent how to use the native tool, UTCP eliminates the symptoms of the middleman:
  \begin{itemize}
    \item<1-> \textbf{No Wrapper Tax:} Just write a simple JSON definition pointing to your existing REST API, gRPC service, or CLI command. No new server needed.
    \vfill
    \item<2-> \textbf{Leverage Existing Infrastructure:} Authentication, permissions, and billing are handled by the tool's native endpoint, as they should be. No re-implementation.
    \vfill
    \item<3-> \textbf{Efficient & Direct:} The agent calls the tool directly. This means lower latency, less overhead, and access to the tool's native, structured data.
  \end{itemize}
\end{frame}

\begin{frame}{Conclusion: Stop Building Middlemen}
  \begin{itemize}
    \item The future of agent-tool integration requires us to move beyond architectures that force developers to build redundant wrappers.
    \vfill
    \item We need protocols that embrace the diversity of existing systems, not replace them.
    \vfill
    \item UTCP provides a path forward: a lean, descriptive standard that makes it easy to connect agents to any tool, on any platform, using the infrastructure that already exists.
  \end{itemize}
\end{frame}

\end{document}
